The persuasion of symbolic elements in government public communications, and their moderation by personal relevance

### 2019-07-12 paper3_relevance and symbols_revision_01.Rmd
Add files via upload

SVIVA2_01_haifa <- filter(SVIVA2_01,AREA==1)
SVIVA2_01_center <- filter(SVIVA2_01,AREA==0)
```

```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(knitr)
library(tidyverse)
library(stargazer)
library(car)
library(broom)
library(kableExtra)
library(ggthemes)
```

@@

# Research design

In this paper, the above-mentioned theoretical expectations are tested via a randomized survey experiment, building on the methodology of @petty_1983. Focusing on the case of the Israeli Environment Protection Ministry (hereafter: EPM), I examine citizens' responses to communications of EPM regarding two of its current policy plans. The first policy regards *reducing the air pollution in the highly polluted Haifa-Bay area*, located in the north of Israel. This policy issue offers a fortunate case for examining the moderating role of perceived personal relevance. The distinction between those who reside in the polluted area and those who do not represents natural variation in personal relevance. Moreover, because this policy issue is intangible, ambiguous and complicated, the differences in personal relevance are not strongly confounded with differences in personal knowledge, as is the case in many other policy issues [@petty_1983]. Congruently, those who reside in the polluted Haifa-Bay area are more motivated to scrutinize the communication about the air-pollution in their area of residence, yet are no more capable of doing so than those who reside elsewhere. As explained below, the sample population of the study consists of residents of the Haifa-bay area, matched with residents of cities at the center of Israel with similar aggregate socio-economic profile. The second policy plan regards *reducing domestic waste and increasing recycling*. This topic is assumed to be equally relevant for Haifa-bay residents and others. Therefore, adding this policy plan is intended to confirm that respondents’ divergent responses to the symbolic elements in the communication regarding the air-pollution policy, if found, may be attributed to differences in perceived personal relevance. I henceforth refer to these two policy plans as "air-pollution" and "recycling".

I exploited three symbolic elements, which have been used in the EPM's public communications, and became widely recognized by the public. The first symbolic element is the unique brand logo of the EPM, which consists a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second element is the green color, which is strongly associated with taking care of the environment. The third element are the images of two high-profile celebrity comedians which were taken from advertising campaigns of the EPM. Further details about these symbolic elements are found in the supplementary appendix.

@@

The survey was conducted online through Qualtrics Survey Software. To capture natural variation in the personal relevance of the air-pollution policy, I sampled citizens from the major cities in the Haifa-Bay area (*Haifa*, *Nesher*, *Qiryat-Hayiim*, *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*), and matched them with citizens from cities with a similar profile and population size at the center of the country (*Natania*, *Petach-Tikva* and *Rishon-Lezion*). To select and recruit residents of these selected cities, I was assisted by an Israeli internet research panel company *iPanel*. The online survey link was sent by iPanel, between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). Importantly, these filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa-bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.

The survey was presented to the participants as a study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. Subjects were also asked to describe the most salient environmental issue in their particular area of residence. These questions were designed to prime subjects to think about the impact of environmental issues to their own lives. Specifically, it was desigend to prime those residing in the Haifa-bay area to think about the personal relevance of air-pollution in their area. Subjects in the relevance control group were asked equivalent questions about their field of occupation.

Participants were then presented with the two policy plans of EPM, and informed that these were obtained from EPM's annual work plan. Each policy plan was followed by questions about respondents' trust in them. Per each policy, subjects were randomly allocated to one of two conditions of substantive information (hereafter: information): strong policy (i.e. logically persuasive policy plan) or weak policy (i.e. logically unpersuasive). Each subject was exposed to two policy plans, one strong and one weak, displayed in a random order. The information in the strong policy plan condition was taken from the EPM’s real work plan and from other official sources, whereas the information in the weak policy plan was fictional, and deliberately designed to create a poor policy plan, which is incompatible with achieving the policy goals. The full texts of the weak and strong conditions for the two policy plans is displayed in **Appendix A**.

The participants were also randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols): treatment ("real symbols"), control-"no symbols" and control-"fake symbols". In the symbol treatment condition, the two policies (as explained, one weak and one strong) are displayed with the EPM logo, green color and celebrity images from campaigns. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities’ images. In the control fake-symbols condition, the EPM logo is replaced by a fake logo, the green color is replaced with blue, and the images of celebrities from the campaigns are replaced with images of unfamiliar people. The symbols conditions are displayed in **Appendix B**. Further details about the real and fake symbols are found in the supplementary appendix.

Altogether, per each policy plan, the experiment has a factorial design of 3 (symbols: real symbols, fake-symbols, no-symbols) x 2 (information: strong vs. weak policy) x 2 (relevance: treatment vs. control). In line with the abovementioned hypotheses, my focus in this study is on the two-way interactions between the two areas and the manipulations of symbols and information. In addition, within the Haifa-Bay respondents, I also explore the two-way interactions between the latter manipulations, and the relevance priming manipulation.[^footnote_information_symbols_paper] I conducted power estimations for the main effects of the symbols and information manipulations, their conditional effects (i.e. within each area) and the aformentioned two-way interactions, based on the analytical sample size. These estimations suggest that the detectible interaction coefficient between areas and the real symbols treatment is 0.27 SDs, and between areas and weak information manipulation is 0.23 SDs. As for the interactions with the relevance manipulation (within the sample of Haifa-Bay respondents), I estimate that their detectible coefficients with the symbols and information conditions are 0.36 and 0.29 SDs, respectively. Full details about these estimations, which were conducted after data collection and analysis, are reported in the Section ## of the supplementary appendix.

[**Tables `r table.x`** about here]

@@

sd.air <- SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()

The outcome variable, citizens' trust in policy, is measured in the survey using a composite index of six items (Cronbach’s alpha = `r alpha_air_trust` and `r alpha_waste_trust` for the air-pollution and waste policies, respectively). This scale was adopted with modifications from validated trust scales used by previous public-administration studies [@karens_2016; @grimmelikhuijsen_2017; @grimmelikhuijsen_2018]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7 (strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan. Participants reported, on average, lower trust in the air-pollution policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`) compared with the recycling policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`).

The key independent variables are the manipulations of symbols and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication.[^footnote_symbols_manipulation_check] The latter is operationalized observationally (through the distinction between the two areas: Haifa-Bay and center) as well as experimentally (through the abovementioned relevance manipulation). I coded living in Haifa-bay area according to respondents’ reporting of residence, work or studying in one of the Haifa-Bay cities.

To confirm the empirical assumption that the air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa-Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents  (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question about the most salient environmental issue in their area of residence, which was presented to the relevance treatment group (i.e. before presenting the policy plans). Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa-Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r (t1/t2*100) %>% round(0)`%).[^footnote_air_relevance_coding] Moreover, many Haifa-Bay respondents explicitly linked the air-pollution to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

@@

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa-Bay respondents' perceptions of personal relevance of the air-pollution policy. On average, respondents in treatment group reported sightly lower level of personal relevance, yet these differences are not statistically significant (*t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

@@

mod_air_1.4 = update(mod_air_1.3,. ~ .+
                      GENDER+
                       GOV_TRUST+
                       IDEOLOGY+
                       EDUCATION+
                       INCOME+
                       CHILDREN+
                       CHILDREN_young)

@@

I begin by exploring respondents trust in the *air-pollution policy*. To reiterate, based on my theoretical hypotheses, I expect that subjects living in the Haifa-Bay, who perceive this issue as personally relevant, would be less affected by the symbolic elements, compared with those living in the Center (H1). Likewise, I expect Haifa-bay respondents to be more affected by the differences in the content (strong versus weak information), compared with center respondents (H2). I first examine the manipulations' main effects (among the entire sample), and then turn to compare their conditional effects (within each of the two areas).

The real symbols condition has a positive and significant main effect on trust in the policy compared with the no symbols condition (*diff* = `r (main.symbols$contrasts %>% tidy() %>% .[1,"estimate"]*-1) %>% max(.) %>% round(.,2)` [`r (t.symbols.no[["conf.int"]][2]*-1) %>% round(.,2)`-`r (t.symbols.no[["conf.int"]][1]*-1) %>% round(.,2)`], *t* = `r t.symbols.no[["statistic"]][["t"]] %>% abs() %>% round(.,3)`). That effect ammounts to `r (main.symbols$contrasts %>% tidy() %>% .[1,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.no[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.no[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`]. Subjects who saw the real symbols also reported, on average, greater trust in the policy, compared with those who received the "fake" symbols, yet these differences (among the entire sample population) were smaller and not sufficiently significant at the 95% level (*diff* = `r (main.symbols$contrasts %>% tidy() %>% .[3,"estimate"]) %>% max(.) %>% round(.,2)` [`r (t.symbols.fake[["conf.int"]][1]) %>% round(.,2)`-`r (t.symbols.fake[["conf.int"]][2]) %>% round(.,2)`], *t* = `r t.symbols.fake[["statistic"]][["t"]] %>% abs() %>% round(.,3)`). The main effect of the weak information condition is negative and significant (*diff* = `r (main.info %>% .[2,"estimate"]*-1) %>% max(.) %>% round(.,2)` [`r (t.info[["conf.int"]][2]) %>% round(.,2)`-`r (t.info[["conf.int"]][1]) %>% round(.,2)`], *t* = `r t.info[["statistic"]][["t"]] %>% abs() %>% round(.,3)`). That effect ammounts to `r ((main.info %>% .[2,"estimate"]*-1)/sd.air) %>% max(.) %>% round(.,2)` SDs [`r (t.info[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`-`r (t.info[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`].

>>>>

SVIVA2_01_haifa <- filter(SVIVA2_01,AREA_center==0)
SVIVA2_01_center <- filter(SVIVA2_01,AREA_center==1)

```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(knitr)
library(tidyverse)
library(stargazer)
library(car)
library(broom)
library(kableExtra)
library(ggthemes)
library(lmtest)
library(sandwich)
```

@@

# Methodology

The above-mentioned theoretical expectations are tested via a randomized survey experiment, building on the methodology of @petty_1983. The experiment focuses, empirically, on the case of the Israeli Environmental Protection Ministry (hereafter: EPM), and its policy for reducing the air pollution in the highly polluted Haifa-Bay area. In the following section, I describe the empirical case, data collection, experimental design and operationalization of research variables.[^footnote_information_symbols_paper]

EPM is a national-level bureau responsible for the formulation, coordination, and execution of environmental policy in Israel. Among others, EPM is responsible for monitoring and regulating the air quality across the country. The Haifa Bay area, located in the north of Israel, has been characterized by relatively high levels of air pollution, mainly due to the concentration of heavy industry in that area, including large oil refineries, power plants, and chemical factories. Beginning in Mid 2000s, EPM has been developing and implementing a series of policy programs to monitor and reduce the high levels of pollution specifically in that area [@theministryofenvironmentprotection_2015; @theministryofenvironmentprotection_2015a; @theministryofenvironmentprotection_2019]. While EPM has claimed that its programs have been effective and led to a significant reduction in air-pollution levels, local residents, professionals and other public agencies have remained highly concerned about the issue, and  accordingly it remained high on the public agenda [@statecomptrollerofisrael_2019]. Specifically, the issue recevied much media and public attention in 2015-2016, following a series of reportings an media publications about the health consequences of the air-pollution in Haifa-Bay. In 2015, the Ministry of Health published statistical data indicating that Haifa-Bay residents experience significanly higher rates of cancer and asthma, compared to other areas of the country. The subsequent media reports specifically highlighted the conclusion that half of the cases of cancer among children in the area can be associated to air-pollution [@elroi_website_2015; @raavad_website_2015]. Thereafter, in 2016, media reporters revealed the interim results of a large epidemiological study, establishing a link between air-pollution in the Haifa-Bay and certain types of cancer, as well as abnormal measurements of newborn babies and growth curves of infants.[^footnote_epidemiological_study] The media reports of that study repeatedly focused on the finding that babys in the Haifa-Bay are born with "smaller heads" [@koriel_website_2016; @rinat_website_2016; @rivlin_website_2016]. These publications, and others, caused much panic among the local population, and especially among young parents.

[^footnote_epidemiological_study]: The epidemiological study was originally sponsored by Ministry of Health. The study, which was  conducted by researchers from Haifa University, was planned to be conducted over five years, between 2015-2020. The findings, which leaked to the media, regarded the interim results of the first year of the study. Following that publication, the study has been significantly criticized by the local research community as well as the professional steering committee of the research. As a result, in 2017, the Ministry decided to cancel its involvement in that research.

The air-pollution in the Haifa-Bay offers a fortunate case for examining the moderating role of perceived personal relevance. The distinction between those who reside in the polluted area and those who do not entails a significant natural variation in that moderating vatiable. Importantly, because this policy issue is intangible and scientifically complicated, the variation in personal relevance is not strongly confounded with personal knowledge about the issue, as is normally the case [@petty_1983]. Accordingly, those residing in the polluted Haifa-Bay area are likely to be more motivated to scrutinize a governemnt communication about the air-pollution in their area, yet they are not necessarily more capable of critically evaluating its content. That latter assumption was the main reason for selecting that case.

While comparing those residing in the Haifa-Bay to others enables me to examine the moderating effect of perceived personal relevance, it still has an important limitation - it represents an observational comparison. The differences in personal relevance can be theoretically confounded by unobserved covariates, that cannot be effectively controlled in the analyses. I address this limitation in the research design in three main ways. *First*, in the selection of the research popultion, I sought to reduce differences in socioeconomic and demographic variables between residents of Haifa-Bay, and their comparison group. Accordingly, I selected the major cities in the Haifa-Bay area (*Haifa*, *Nesher*, *Qiryat-Hayiim*, *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*), and matched them with citizens from cities with a similar profile and population size at the center of the country (*Natania*, *Petach-Tikva* and *Rishon-Lezion*). I also searched EPM documents and news websites for any reporting about high air-pollution levels in the three center cities, to ensure that the two groups would differ in relation to the their perceptions about the relevance air-pollution policies. *Second*, in addition to the policy regarding the air-pollution in Haifa-Bay, I also included in the experiment a second policy plan that is expected to be percevied by the two groups as equally relevant. For this purpose, I selected a policy plan of EPM that regards *reducing domestic waste and increasing recycling*. The comparison between the two EPM policies is intended to confirm that respondents’ divergent responses to symbolic elements and variation in substantive information in the communication of the first policy, if found, are attributed to differences in perceived personal relevance, and not to other covariates. I henceforth refer to these two policy plans as "air-pollution" and "recycling". *Third*, in addition to the observational comparison between the two areas, I also included an experimental manipulation aimed at enhancing the perceived personal relevance of the air-pollution policy among the Haifa-Bay residents.

@@

To select and recruit residents of these selected cities, I was assisted by an Israeli internet research panel company *iPanel*. The survey experiment was conducted online through Qualtrics Survey Software. The online survey link was sent by iPanel between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). Thereafter, I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). These filters are neither associated with the areas nor with the experimental conditions. Following this screening, I ended up with a sizable analytical sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa-bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.[^footnote_unfiltered_sample]

[^footnote_unfiltered_sample]: In the supplementary appendix, I replicate the main models on the unfiltered sample, with no significant change to the results.

The survey was presented to the participants as an academic study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, all subjects, regardless of their area of residence, were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance manipulation). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. Then, they were asked to describe the most salient environmental issue in their area of residence. These treatment questions were designed to prime those residing in the Haifa-bay area to think about the personal relevance of air-pollution in their area.[^footnote_relevance_manipulation] Subjects in the relevance control group were asked equivalent questions about their field of occupation.

[^footnote_relevance_manipulation]: The first three treatment questions were also designed to prime subjects, more generally, to think about the impact of environmental issues to their own lives. In this paper, I focus on the the moderating effect of the relevance manipulation on Haifa-Bay citizens' responses to the air-pollution policy. In other analyses, presented in the supplementary appendix, I also examine the effect of that variable for the recycling policy, for both populations.[##ADD##]

Participants were then presented with the air-pollution and recycling policy plans (displayed in a random order). They were informed that these plans were obtained from EPM's annual work plan. Per each policy, subjects were randomly allocated to one of two conditions of substantive information (hereafter: information manipulation): strong policy (i.e. logically persuasive policy plan) or weak policy (i.e. logically unpersuasive). Each subject was exposed to two policy plans, one strong and one weak, displayed in a random order. For instance, a subject who first received a strong air-pollution policy plan, later received the weak recycling policy plan, and vice versa. The information in the strong policy plan condition was taken from the EPM’s real work plan and from other official sources, whereas the information in the weak policy plan was fictional, and deliberately designed to create a poor policy plan, which is incompatible with achieving the policy goals. The full texts of the weak and strong conditions for the two policy plans is displayed in **Appendix A**.

The participants were also randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols manipulation): treatment ("real symbols"), control-"no symbols" and control-"fake symbols". In the real symbols condition, the two policies (as explained, one weak and one strong) are displayed with three symbolic elements, which have been used in EPM's public communications, and became widely recognized by the public. The first symbolic element is the unique brand logo of the EPM, which consists a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second element is the green color, which is strongly associated with taking care of the environment. The third element is an image of a celebrity comedian from an advertising campaign of the EPM. The air-pollution policy (in the real symbols condition) was presented with an image from EPM campaign from 2010 under the slogan "starting to think green", starring the Israeli celebrity comedian Tal Friedman. The recycling policy in the real symbols condition was presented with an image from a campaign from 2017 for reducing the use of disposable bags ("taking every bag seriously"), starring the comedian Ido Rosenblum. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities’ images. In the control fake-symbols condition, the EPM logo is replaced by a fake logo, the green color is replaced with blue, and the images of celebrities from the campaigns are replaced with edited images of unfamiliar people. The purpose of the fake-logo condition was to distinguish the effect of the symbolic meaning of the EPM's graphic identifiers (i.e. the symbolic effect) from a possible effect of their aesthetic design. Accordingly, I specifically designed the "fake" symbols so that they would resemble the real symbols and have equivalent aesthetic qualities, without activating particular emotions and associations related to EPM, or environmental policy. The symbols conditions are displayed in **Appendix B**. Further details about the real and fake symbols are found in section ## of the supplementary appendix.

Each policy plan was followed by six questions about respondents' trust in the policy. After the trust questions of the second policy, the subjects were asked additional questions about the two policies, including the degree to which they perceive each of them as personally relevant. These questions were followed by an Instructional Manipulation Check (IMC) item, intended to assess participants’ diligence. The last sections included a set of demographic questions including questions about their place of residence and work, and manipulation checks for participants' familiarity with the real symbols. At the end of the survey, the participants were debriefed about the manipulations, and the possible deception. An English translation of the full survey is available in section ## of the supplementary Appendix.

In line with the abovementioned hypotheses, my main focus in this paper is on participants' trust in the air-pollution policy, and the two-way interactions between residence in Haifa-Bay and the symbols and information manipulations of that policy. I also explore these interactions with regard to the recycling policy, and compare them to the air-pollution interactions. Additionally, within the Haifa-Bay sample, I examine the two-way interactions between the relevance manipulation and the symbols and information manipulations. I conducted power estimations for the main effects of the symbols and information manipulations, their conditional effects within each area and the relevant two-way interactions. I conducted these estimations after the data collection and analyses, based on the analytical sample size. The estimated detectible interaction coefficient between areas and the real symbols treatment is 0.27 SDs, and between areas and weak information manipulation is 0.23 SDs. As for the interactions with the relevance manipulation (among Haifa-Bay respondents), I estimate that the detectible coefficients of their interactions with the symbols and information conditions are 0.36 and 0.29 SDs, respectively. The power estimations are reported in details in Section ## of the supplementary appendix.

@@

sd.air <- SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()
sd.air.haifa <- SVIVA2_01_comb %>% filter(AREA_center==0)%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()
```

The outcome variable, citizens' trust in policy, is measured in the survey using a composite index of six items (Cronbach’s alpha = `r alpha_air_trust` for air-pollution policy, and `r alpha_waste_trust` for the recycling policy). This scale was adopted with modifications from validated trust scales used by previous public-administration studies [@karens_2016; @grimmelikhuijsen_2017; @grimmelikhuijsen_2018]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7 (strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan.

The key independent variables are the manipulations of symbols[^footnote_symbols_manipulation_check] and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication. The latter is operationalized observationally (through the distinction between the two areas: Haifa-Bay and center) as well as experimentally (through the abovementioned relevance manipulation).

Living in Haifa-bay area was coded according to respondents’ reporting of residence, work or studying in one of the Haifa-Bay cities. To confirm the empirical assumption that the air-pollution policy (but not the recycling policy) is perceived as more personally relevant by Haifa Bay residents compared with center residents, respondents were asked, after the outcome variable items: "To what extent did you feel that the policy on ["Reducing the air-pollution in the Haifa-Bay" / “reducing waste and increasing recycling”] is personally relevant to you?" (1="very little"; 7="very much"). `r figure.x=figure.x+1`**Figure `r figure.x`** shows the distributions of the latter items across the two areas. As expected, the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents  (Cohen's D = `r relevance.areas.d[["estimate"]] %>% round(2)` [`r relevance.areas.d[["conf.int"]][["lower"]] %>% round(2)`,`r relevance.areas.d[["conf.int"]][["upper"]] %>% round(2)`]). By comparison, the recycling policy is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them. Additional support for these differences are provided by respondents' answers to the open question about the most salient environmental issue in their area of residence, which was presented to the relevance treatment group (i.e. before presenting the policy plans). Indeed, `r t1<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow()``r t1` of the `r t2<-SVIVA2_01_haifa %>% filter(RELEVANCE_exp==1) %>% nrow()``r t2` Haifa-Bay subjects who were asked this question regarded in their responses to the air-pollution generated from the industrial area (`r (t1/t2*100) %>% round(0)`%).[^footnote_air_relevance_coding] Moreover, many Haifa-Bay respondents explicitly linked the air-pollution to higher morbidity rates in their area. Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, and only `r ((SVIVA2_01_center %>% filter(RELEVANCE_exp==1,air.content==1) %>% nrow())/nrow(SVIVA2_01_center)*100) %>% round(0)`% of them mentioned air-pollution as a prominent environmental concern in their area.

@@

As for the relevance manipulation, data suggests that assignment to relevance priming questions did not increase Haifa-Bay respondents' perceptions of personal relevance of the air-pollution policy. In fact, those assigned to the treatment group reported sightly lower level of personal relevance on average, yet these differences are not statistically significant (*t* = `r relevance.exp.t[["statistic"]] %>% round(3)`, *p*=`r round(relevance.exp.t[["p.value"]],3)`). This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

@@

mod_air_1.4 = update(mod_air_1.3,. ~ .+
                      GENDER+
                       AGE+
                       GOV_TRUST+
                       IDEOLOGY+
                       EDUCATION+
                       INCOME+
                       CHILDREN_young)

@@

I now turn to exploring respondents trust in the *air-pollution policy*. To reiterate, based on my theoretical hypotheses, I expect that subjects living in the Haifa-Bay, who perceive this issue as personally relevant, would be less affected by the symbolic elements, compared with those living in the Center (H1). Likewise, I expect Haifa-bay respondents to be more affected by the differences in the content (strong versus weak information), compared with center respondents (H2).

This outcome variable has a right-skewed distribution (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`, *Median* = `r SVIVA2_01_comb%>%summarise(median(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`), whereby more subjects reported very low levels of trust, than very high levels of trust in than policy. The real symbols condition has a positive and significant main effect on trust in that policy plan compared with the no symbols condition (*diff* = `r (main.symbols$contrasts %>% tidy() %>% .[1,"estimate"]*-1) %>% max(.) %>% round(.,2)` [`r (t.symbols.no[["conf.int"]][2]*-1) %>% round(.,2)`-`r (t.symbols.no[["conf.int"]][1]*-1) %>% round(.,2)`], *t* = `r t.symbols.no[["statistic"]][["t"]] %>% abs() %>% round(.,3)`). That effect ammounts to `r (main.symbols$contrasts %>% tidy() %>% .[1,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.no[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.no[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`]. Subjects who saw the real symbols also reported, on average, greater trust in the policy, compared with those who received fake symbols, yet these differences are not significant at the 95% level (*diff* = `r (main.symbols$contrasts %>% tidy() %>% .[3,"estimate"]) %>% max(.) %>% round(.,2)` [`r (t.symbols.fake[["conf.int"]][1]) %>% round(.,2)`-`r (t.symbols.fake[["conf.int"]][2]) %>% round(.,2)`], *t* = `r t.symbols.fake[["statistic"]][["t"]] %>% abs() %>% round(.,3)`, *p* = `r (t.symbols.fake[["p.value"]]/2) %>% abs() %>% round(.,3)`, one-tailed test). The main effect of the weak information condition is negative and significant (*diff* = `r (main.info %>% .[2,"estimate"]*-1) %>% max(.) %>% round(.,2)` [`r (t.info[["conf.int"]][2]) %>% round(.,2)`-`r (t.info[["conf.int"]][1]) %>% round(.,2)`], *t* = `r t.info[["statistic"]][["t"]] %>% abs() %>% round(.,3)`). That effect ammounts to `r ((main.info %>% .[2,"estimate"]*-1)/sd.air) %>% max(.) %>% round(.,2)` SDs [`r (t.info[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`-`r (t.info[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`].

`r figure.x=figure.x+1`**Figure `r figure.x`** graphically illustrates the conditional effects of the manipulations, within each of the two areas.[^footnote_raw_means] As shown by the figure, participants from Haifa-Bay reported significanly lower levels of trust in the air-pollution policy, in all the experimental conditions. The lower intercept of Haifa-Bay respondents suggests that they tend to have prior negative attitudes towards EPM's actions regarding the air-pollution in their area. As for the effect of the symbols manipulation (presented in the left panel), the results show that the symbols had a greater effect on those living in the polluted Haifa-Bay area. Their exposure to the real symbols increased their trust in the policy by `r (cond.symbols.haifa$contrasts %>% tidy() %>% .[1,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.no.haifa[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.no.haifa[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`] comapred with the no symbols group, and `r (cond.symbols.haifa$contrasts %>% tidy() %>% .[3,"estimate"] %>% max(.)/sd.air) %>% abs() %>% round(2)` SDs [`r (t.symbols.fake.haifa[["conf.int"]][1]/sd.air) %>% abs() %>% round(2)`-`r (t.symbols.fake.haifa[["conf.int"]][2]/sd.air) %>% abs() %>% round(2)`] comapred to the fake symbols group. The center participants who saw the real symbols also reported, on average, higher levels of trust in the policy, compared with the two control groups. However, these differences were smaller compared with the effect among the Haifa-Bay subjects, and not sufficiently significant. The effects of the information manipulation on the two groups are presented in the right panel. This comparison shows that the differences bewteen participants' trust in the weak and strong policies were roughly similar in both groups.

[^footnote_raw_means]: In addition, raw means and SDs of outcome variables across the two areas are reported in section XX of the supplementary appendix.
